CRUSE to Safe Cycling in Ireland \\\large{An Open Source Methodology to Support Active Travel}

### 2023-09-22 paper.qmd
Add shiny citation

# Uncomment for draft versions:
format: 
  html: default
  docx: default

# # Uncomment for formatted version
# format:
#   elsevier-pdf:
#     keep-tex: true
#     journal:
#       name: Journal Name
#       formatting: preprint
#       model: 3p
#       cite-style: super

@@

A proactive approach to the development of cycling policy and
infrastructure is taking place in Ireland.
The main framework
underpinning these efforts is the Climate Action Plan[^2] and The National
Development Plan[^3]
. Strategic alignment is also outlined by the
Phoenix Park Transport and Mobility Options Study, which highlights public support for active travel interventions: \"83% of respondents supported enhanced walking and cycling facilities\"
.

@@

At the regional level, the recently published Greater Dublin Area Transport Strategy[^4] reinforces these findings: \"nearly a quarter of adults cycle at least once a week in the Dublin Metropolitan Area\" with cycling in the Dublin area taking up to 60,000 cars off the road today.
Extrapolating this on a per population basis across Ireland, with around 40% of the population living in Dublin, this suggests that around 150,000 cars could be removed nationwide just by achieving Dublin levels of cycling in all counties (notwithstanding existing cycling trips and differences in trip distances).

@@

It was in this context that Transport Infrastructure Ireland (TII) commissioned the Cycle Route Uptake and Scenario Estimation (CRUSE) project.
Based on the Propensity to Cycle Tool (PCT) for England and Wales, the CRUSE tool was developed to provide evidence on current cycling levels and future cycling potential nationwide across Ireland.
A key consideration for TII is to ensure both urban and rural Ireland were targeted for cycling infrastructure.
To this end, TII extensively reviewed activity on Irish Roads and how the road network was being used by commuters, cyclists, heavy goods vehicles and other road users across the Irish transport network.
The aim was for the tool to provide strong, national, systematic but locally-specific evidence to monitor cycling friendliness and safety in Ireland, to ensure strategic alignment with national, regional and local policies.
The open source and publicly available nature of the tool ensures more inclusive and evidence-based conversations around cycle network planning between all stakeholders in the transport planning process.
Other use cases in the Irish context includes monitoring the need for existing cycling infrastructure upgrades, undertaking appraisal, as well as safety and exposure information in line with the European Cycling Federation, Dutch Cycling Embassy recommendations and the European RISM Directive, for reporting collision rates of vulnerable road users by 2024.
These factors mean that the tool can be seen as an open access 'leverage point' in the planning system [@lovelace2020].

@@

However, travel to school and work only constitute a small proportion of all trips.
To overcome this issue, we developed a spatial interaction modelling methodology to estimate the number of trips between each OD pair for additional trip purposes.

The classification of trip purposes used in the CRUSE tool was guided mainly by the trip purpose classification found within the National Household Travel Survey (NHTS), but with the addition of categories based on the comprehensive POWSCAR data, and the need to include recreational trips and multi-stage trips (not yet implemented).
An overview of the trip purposes used in CRUSE is presented in @fig-trip-purposes.

@@

>>>>

# # Uncomment for draft versions:
# format: 
#   html: default
#   docx: default

# Uncomment for formatted version
format:
  elsevier-pdf:
    keep-tex: true
    journal:
      name: Journal Name
      formatting: preprint
      model: 3p
      cite-style: super

@@

A proactive approach to the development of cycling policy and infrastructure is taking place in Ireland.
The main framework underpinning these efforts is the Climate Action Plan[^2] and The National Development Plan[^3]
. Strategic alignment is also outlined by the Phoenix Park Transport and Mobility Options Study, which highlights public support for active travel interventions: "83% of respondents supported enhanced walking and cycling facilities"
.

@@

At the regional level, the recently published Greater Dublin Area Transport Strategy[^4] reinforces these findings: "nearly a quarter of adults cycle at least once a week in the Dublin Metropolitan Area" with cycling in the Dublin area taking up to 60,000 cars off the road today.
Extrapolating this on a per population basis across Ireland, with around 40% of the population living in Dublin, this suggests that around 150,000 cars could be removed nationwide just by achieving Dublin levels of cycling in all counties (notwithstanding existing cycling trips and differences in trip distances).

@@

It was in this context that Transport Infrastructure Ireland (TII) commissioned the Cycle Route Uptake and Scenario Estimation (CRUSE) project.
Based on the Propensity to Cycle Tool (PCT) for England and Wales, the CRUSE tool was developed to provide evidence on current cycling levels and future cycling potential nationwide across Ireland.
A key consideration for TII is to ensure both urban and rural Ireland were targeted for cycling infrastructure, as shown in @fig-dublin.
To this end, TII extensively reviewed activity on Irish Roads and how the road network was being used by commuters, cyclists, heavy goods vehicles and other road users across the Irish transport network.
The aim was for the tool to provide strong, national, systematic but locally-specific evidence to monitor cycling friendliness and safety in Ireland, to ensure strategic alignment with national, regional and local policies.
The open source and publicly available nature of the tool ensures more inclusive and evidence-based conversations around cycle network planning between all stakeholders in the transport planning process.
Other use cases in the Irish context includes monitoring the need for existing cycling infrastructure upgrades, undertaking appraisal, as well as safety and exposure information in line with the European Cycling Federation, Dutch Cycling Embassy recommendations and the European RISM Directive, for reporting collision rates of vulnerable road users by 2024.
These factors mean that the tool can be seen as an open access 'leverage point' in the planning system [@lovelace2020].

@@

However, travel to school and work only constitute a small proportion of all trips.
To overcome this issue, we developed a spatial interaction modelling methodology to estimate the number of trips between each OD pair for additional trip purposes.
The classification of trip purposes used in the CRUSE tool was guided mainly by the trip purpose classification found within the National Household Travel Survey (NHTS), but with the addition of categories based on the comprehensive POWSCAR data, and the need to include recreational trips and multi-stage trips (not yet implemented).
An overview of the trip purposes used in CRUSE is presented in @fig-trip-purposes.

@@

The CRUSE web application is statically hosted, meaning that it can be accessed without the need for a server running software such as the R package 'shiny' or Python packages such as 'streamlit' or 'flask' in the background [@wickham2021].

### 2023-09-22 paper.tex
Add shiny citation

\makeatletter
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother
\makeatletter
\@ifpackageloaded{tcolorbox}{}{\usepackage[skins,breakable]{tcolorbox}}
\makeatother
\makeatletter
\@ifundefined{shadecolor}{\definecolor{shadecolor}{rgb}{.97, .97, .97}}
\makeatother
\makeatletter
\makeatother
\makeatletter
\makeatother
\journal{Journal Name}
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\usepackage[]{natbib}
\bibliographystyle{elsarticle-num}
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={CRUSE to Safe Cycling in Ireland},
  pdfauthor={Dr.~Robin Lovelace; Dr.~Joey Talbot; Eugeni Vidal Tortosa; Hussein Mahfouz; Elaine Brick; Peter Wright; Gary O'Tool; Dan Brennan; Dr.~Suzanne Meade},
  pdfkeywords={Cycling, Open-source, Road Safety, Active Travel},
  colorlinks=true,
  linkcolor={blue},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}

@@

\begin{frontmatter}
\title{CRUSE to Safe Cycling in Ireland \\\large{An Open Source
Methodology to Support Active Travel} }
\author[1]{Dr.~Robin Lovelace%
\corref{cor1}%
}
 \ead{r.lovelace@leeds.ac.uk} 
\author[1]{Dr.~Joey Talbot%
%
}

@@

\author[3]{Gary O'Tool%
%
}

@@

\author[4]{Dr.~Suzanne Meade%
%
}

@@

\begin{keyword}
    Cycling \sep Open-source \sep Road Safety \sep 
    Active Travel
\end{keyword}
\end{frontmatter}
    \ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[boxrule=0pt, borderline west={3pt}{0pt}{shadecolor}, interior hidden, breakable, frame hidden, sharp corners, enhanced]}{\end{tcolorbox}}\fi

@@

It was in this context that Transport Infrastructure Ireland (TII)
commissioned the Cycle Route Uptake and Scenario Estimation (CRUSE)
project. Based on the Propensity to Cycle Tool (PCT) for England and
Wales \citep{lovelace2017}, the CRUSE tool was developed to provide
evidence on current cycling levels and future cycling potential
nationwide across Ireland.

@@

The methods used to generate the evidence presented in the CRUSE tool
for Ireland build on the Propensity to Cycle Tool (PCT), which was
originally funded by the UK's Department for Transport and developed by
a multi-university team. An important feature of the PCT is that it is
open source and publicly available (at
\href{https://www.pct.bike/}{www.pct.bike}), allowing its use by all
stakeholders in the transport planning process \citep{lovelace2017}. The
PCT approach has had major policy impacts, as outlined in Research
Excellence Framework (REF) impact case studies, which demonstrate that
the tool ``revolutionised strategic cycle planning in England and
Wales''\footnote{See REF Impact Case Study ``Cycle network policy,
  planning and investment transformed by the Propensity to Cycle Tool''
  at
  \href{https://results2021.ref.ac.uk/impact/847d1191-7f25-46ba-a399-b481125edc8f}{results2021.ref.ac.uk}
  submitted by the University of Leeds.} by overcoming the barriers to
cycling investment imposed by lack of evidence on cycling
potential\footnote{See the REF Impact Case Study ``Creating Step Changes
  in Cycling Policy and Infrastructure Planning across the UK'' at
  \href{https://results2021.ref.ac.uk/impact/4BBF3436-FD10-4C75-9791-F5E98AB4411B}{results2021.ref.ac.uk}
  submitted by the University of Westminster.} .

@@

@@

@@

@@

bla bla

>>>>

\makeatletter
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother
\makeatletter
\@ifpackageloaded{tcolorbox}{}{\usepackage[skins,breakable]{tcolorbox}}
\makeatother
\makeatletter
\@ifundefined{shadecolor}{\definecolor{shadecolor}{rgb}{.97, .97, .97}}
\makeatother
\makeatletter
\makeatother
\makeatletter
\makeatother
\journal{Journal Name}
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\usepackage[]{natbib}
\bibliographystyle{elsarticle-num}
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={CRUSE to Safe Cycling in Ireland},
  pdfauthor={Robin Lovelace; Joey Talbot; Eugeni Vidal Tortosa; Hussein Mahfouz; Elaine Brick; Peter Wright; Gary O'Tool; Dan Brennan; Suzanne Meade},
  pdfkeywords={Cycling, Open-source, Road Safety, Active Travel},
  colorlinks=true,
  linkcolor={blue},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}

@@

\begin{frontmatter}
\title{CRUSE to Safe Cycling in Ireland \\\large{An Open Source
Methodology to Support Active Travel} }
\author[1]{Robin Lovelace%
\corref{cor1}%
}
 \ead{r.lovelace@leeds.ac.uk} 
\author[1]{Joey Talbot%
%
}

@@

\author[2]{Gary O'Tool%
%
}

@@

\author[4]{Suzanne Meade%
%
}

@@

\begin{keyword}
    Cycling \sep Open-source \sep Road Safety \sep 
    Active Travel
\end{keyword}
\end{frontmatter}
    \ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[frame hidden, boxrule=0pt, interior hidden, sharp corners, borderline west={3pt}{0pt}{shadecolor}, enhanced, breakable]}{\end{tcolorbox}}\fi

@@

A proactive approach to the development of cycling policy and
infrastructure is taking place in Ireland. The main framework
underpinning these efforts is the Climate Action Plan\footnote{\url{https://www.gov.ie/en/publication/7bd8c-climate-action-plan-2023/}}
and The National Development Plan\footnote{\url{https://www.gov.ie/en/publication/774e2-national-development-plan-2021-2030/}}
. Strategic alignment is also outlined by the Phoenix Park Transport and
Mobility Options Study, which highlights public support for active
travel interventions: ``83\% of respondents supported enhanced walking
and cycling facilities'' .

At the regional level, the recently published Greater Dublin Area
Transport Strategy\footnote{\url{https://www.nationaltransport.ie/planning-and-investment/strategic-planning/greater-dublin-area-transport-strategy/}}
reinforces these findings: ``nearly a quarter of adults cycle at least
once a week in the Dublin Metropolitan Area'' with cycling in the Dublin
area taking up to 60,000 cars off the road today. Extrapolating this on
a per population basis across Ireland, with around 40\% of the
population living in Dublin, this suggests that around 150,000 cars
could be removed nationwide just by achieving Dublin levels of cycling
in all counties (notwithstanding existing cycling trips and differences
in trip distances).

Further evidence of the importance of cycling in Ireland \emph{already}
is provided by the National Strategic Objective (NSO) from the National
Development Plan, with €8.6 billion allocated to sustainable transport
infrastructure including public transport and active travel
interventions. Cycle infrastructure will be developed in synchrony with
the BusConnects project, an entire redesign of the bus network in Dublin
and Cork.

It was in this context that Transport Infrastructure Ireland (TII)
commissioned the Cycle Route Uptake and Scenario Estimation (CRUSE)
project. Based on the Propensity to Cycle Tool (PCT) for England and
Wales, the CRUSE tool was developed to provide evidence on current
cycling levels and future cycling potential nationwide across Ireland. A
key consideration for TII is to ensure both urban and rural Ireland were
targeted for cycling infrastructure, as shown in
Figure~\ref{fig-dublin}. To this end, TII extensively reviewed activity
on Irish Roads and how the road network was being used by commuters,
cyclists, heavy goods vehicles and other road users across the Irish
transport network. The aim was for the tool to provide strong, national,
systematic but locally-specific evidence to monitor cycling friendliness
and safety in Ireland, to ensure strategic alignment with national,
regional and local policies. The open source and publicly available
nature of the tool ensures more inclusive and evidence-based
conversations around cycle network planning between all stakeholders in
the transport planning process. Other use cases in the Irish context
includes monitoring the need for existing cycling infrastructure
upgrades, undertaking appraisal, as well as safety and exposure
information in line with the European Cycling Federation, Dutch Cycling
Embassy recommendations and the European RISM Directive, for reporting
collision rates of vulnerable road users by 2024. These factors mean
that the tool can be seen as an open access `leverage point' in the
planning system \citep{lovelace2020}.

@@

The methods used to generate the evidence presented in the CRUSE tool
for Ireland build on the Propensity to Cycle Tool (PCT), which was
originally funded by the UK's Department for Transport and developed by
a multi-university team. An important feature of the PCT is that it is
open source and publicly available (at
\href{https://www.pct.bike/}{www.pct.bike}), allowing its use by all
stakeholders in the transport planning process \citep{lovelace2017}. The
PCT approach has had major policy impacts, as outlined in Research
Excellence Framework (REF) impact case studies, which demonstrate that
the tool ``revolutionised strategic cycle planning in England and
Wales''\footnote{See REF Impact Case Study ``Cycle network policy,
  planning and investment transformed by the Propensity to Cycle Tool''
  at
  \href{https://results2021.ref.ac.uk/impact/847d1191-7f25-46ba-a399-b481125edc8f}{results2021.ref.ac.uk}
  submitted by the University of Leeds.} by overcoming the barriers to
cycling investment imposed by lack of evidence on cycling
potential\footnote{See the REF Impact Case Study ``Creating Step Changes
  in Cycling Policy and Infrastructure Planning across the UK'' at
  \href{https://results2021.ref.ac.uk/impact/4BBF3436-FD10-4C75-9791-F5E98AB4411B}{results2021.ref.ac.uk}
  submitted by the University of Westminster.}.

@@

A limitation of the original PCT was that it only included travel to
work data. This was partially addressed by the inclusion of travel to
school based on data from the Department for Education in England
\citep{goodman2019}. An advantage of the POWSCAR OD data over OD
datasets derived from census surveys in many countries is that it
includes travel to school data. We commissioned a version of POWSCAR
that included a breakdown of the total flows between OD pairs by purpose
and mode, enabling a more realistic estimation of the `Baseline' cycling
network.

However, travel to school and work only constitute a small proportion of
all trips. To overcome this issue, we developed a spatial interaction
modelling methodology to estimate the number of trips between each OD
pair for additional trip purposes.

The classification of trip purposes used in the CRUSE tool was guided
mainly by the trip purpose classification found within the National
Household Travel Survey (NHTS), but with the addition of categories
based on the comprehensive POWSCAR data, and the need to include
recreational trips and multi-stage trips (not yet implemented). An
overview of the trip purposes used in CRUSE is presented in
Figure~\ref{fig-trip-purposes}.

@@

\includegraphics{images/paste-4.png}

@@

\caption{\label{fig-trip-purposes}Trip purposes used in the CRUSE tool.}

@@
